Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra using Murchison Widefield Array Observations

### 2015-04-08 ms.tex
Modify Section 5
From e2e1f01:
============

In ms.tex,

(*) Modify Section 5

Sensitive instruments such as the Square Kilometre Array (SKA) with the capability of direct imaging of redshifted H~{\sc i} are yet to become operational. In the meanwhile, the Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) currently under construction will be much closer in its capability to detect and place definitive constraints on the reionization epoch relative to the SKA and far ahead of current instruments such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, which have enough sensitivity for a statistical detection of the signal \citep{bow06,par12a,bea13,dil13,thy13,pob14}.

@@

Fluctuations of redshifted H~{\sc i} from the reionization epoch are extremely faint relative to the radio foregrounds. This poses perhaps the greatest challenge to EoR experiments and thus merits a detailed characterization in order to separate them from the desired cosmological H~{\sc i} signal.

In a recent study, through the use of all-sky foreground and instrument models, the role of wide-field nature of radio interferometric measurements, hitherto unpredicted, has emerged. Referred to as the {\it pitchfork} signature in the delay spectral domain, it is characterized by significant foreground power in the primary field of view and an enhancement in power near the horizon. The latter is due to the highly nonlinear mapping between geometric delays and subtended solid angles, besides an increase in sensitivity to larger scales caused by foreshortening of baseline lengths along these directions. Owing to its localization in the delay domain, it is a leading cause of contamination in Fourier modes most sensitive to the EoR H~{\sc i} signal.

Knowledge of these characteristics weigh heavily on choices in instrument design like antenna aperture and in data analysis such as optimal weighting schemes for future instruments such as HERA and SKA. We report the first robust detection of the {\it pitchfork} signature (which could not be confirmed earlier in single MWA snapshots due to large thermal fluctuations) with a signal-noise ratio $>10$ made possible by coherent averaging of 12 independent LST-aligned snapshots from MWA observations.

% By averaging 12 independent LST-aligned snapshots from MWA observations, we improve the dynamic range in delay power spectra by a factor $\gtrsim$~10 while maintaining the foreground contributions in the observed {\it visibilities} coherent. Foreground power in delay bins near the horizon is distinctly visible with a signal-noise ratio $>10$. Further, there is a close resemblance in the morphologies of delay power spectra from data and models. Together, they allow us to make the first robust confirmation of the nature of wide-field measurements in EoR experiments.

>>>>

Sensitive instruments such as the Square Kilometre Array (SKA) with the capability of direct imaging of redshifted H~{\sc i} are yet to become operational. In the meanwhile, the Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA), currently under construction, will be much closer in its capability to detect and place definitive constraints on the reionization epoch relative to the SKA and far ahead of current instruments such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, which have enough sensitivity for a statistical detection of the signal \citep{bow06,par12a,bea13,dil13,thy13,pob14}.

@@

Using deeper MWA data, we have confirmed with high significance the earlier prediction that wide-field EoR measurements suffer significant foreground contamination from near the horizon. This has important implications for instrument design and data analysis of future instruments such as HERA and SKA.

Precise modeling is thus required to gain a complete understanding of the characteristics of the cosmological signal and the foregrounds. All-sky surveys matching the frequency and angular resolution of observation, and a detailed model of the instrument will further improve the accuracy of our models. In our earlier study, we proposed a selective flagging of data on different antenna spacings that can potentially mitigate foreground contamination by two orders of magnitude. Efforts are underway to incorporate this proposed foreground mitigation technique into the MWA data analysis algorithms.

For future work, we plan to extend our analysis to the upcoming instrument, HERA, which is currently under construction. It is a closely packed hexagonal array of fixed 14~m dishes which will observe the sky drifting overhead with redundant antenna spacings. Based on our earlier study, such a dish will have a much desirable Fourier response from a foreground contamination viewpoint. One of our objectives is to forecast the per-baseline foreground contamination as a function of local sidereal time in order to tune the observing strategy to maximize sensitivity to the cosmological signal.

% Fluctuations of redshifted H~{\sc i} from the reionization epoch are extremely faint relative to the radio foregrounds. This poses perhaps the greatest challenge to EoR experiments and thus merits a detailed characterization in order to separate them from the desired cosmological H~{\sc i} signal.

% In a recent study, through the use of all-sky foreground and instrument models, the role of wide-field nature of radio interferometric measurements, hitherto unpredicted, has emerged. Referred to as the {\it pitchfork} signature in the delay spectral domain, it is characterized by significant foreground power in the primary field of view and an enhancement in power near the horizon. The latter is due to the highly nonlinear mapping between geometric delays and subtended solid angles, besides an increase in sensitivity to larger scales caused by foreshortening of baseline lengths along these directions. Owing to its localization in the delay domain, it is a leading cause of contamination in Fourier modes most sensitive to the EoR H~{\sc i} signal.

% Knowledge of these characteristics weigh heavily on choices in instrument design like antenna aperture and in data analysis such as optimal weighting schemes for future instruments such as HERA and SKA. We report the first robust detection of the {\it pitchfork} signature (which could not be confirmed earlier in single MWA snapshots due to large thermal fluctuations) with a signal-noise ratio $>10$ made possible by coherent averaging of 12 independent LST-aligned snapshots from MWA observations.
